---
title: "Project"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code
## Setup the data

First we load the data from the CSV published on MSU website, we then perform
some pre-processing to create counties_list and top_ten_counties_list. counties_list
is a vector of length $n$ which is the number of samples. Each element in the vector
is just a number indicating the county number. 
For example, if Gallatin County has 2,134 students at MSU and is county 6, then there will 
be 2,134 6's in the counties_list vector. top_ten_counties_list is just the counties_list
subsetted to only contain the 10 most represented counties at MSU

```{r}
library(readr)
counties <- read_csv("county.csv")
counties = as.matrix(counties)[-c(59,58,57),] #get rid of row of NAs, unspecified counties, and total
top_ten_counties <- counties[,1][order(counties[,2])][47:56]
top_ten_counties <- rev(top_ten_counties)

##All counties
counties_list <- c()
for(i in 1:dim(counties)[1]){
  counties_list <- append(counties_list, 
                          rep(i,
                              times = counties[i,2])) #counties_matrix[i,1] 
}
##Top 10 counties
top_ten_counties_list <- c()
for(i in 1:length(top_ten_counties)){
  top_ten_counties_list <- append(top_ten_counties_list,
                                  rep(i,
                                      counties[,2][which(counties[,1]==top_ten_counties[i])]))
}
```

## Functions
We declare some helper functions. dist computes the similarity between the two probabilities (histograms).
We use the KL-Divergence of the bootstrap sample (P) to the true sample (Q) (KL-divergence is non-symmetric).
We then also declare the function run_bootstraps that runs our experiments.
```{r}
dist <- function(P, Q){  # chisquare distance
  return ((1/2)  * sum( (P - Q)^2 / (P + Q + 0.000000001)^2 ))  # add small number so not divide by zero
  #return (sum(P * log(P/Q))) # KL-Divergence
}


run_bootstraps <- function(reps, counties){
  library(pracma)
  library(purrr)
  n_list = c(50, 100, 500, 1000)
  dists <- rep(0,times=length(n_list)*reps)
  counter = 1
  
  county_counts = unlist(map(seq.int(0, 57), ~ sum(counties == .x)))  / length(counties)
  
  for(i in 1:reps){
    for(n in n_list){
      our_sample <- sample(counties,30)
      bootstrap_pop <- sample(our_sample, n, replace = TRUE)
      bootstrap_counts = unlist(map(seq.int(0, 57), ~ sum(bootstrap_pop == .x)))/length(bootstrap_pop)
      dists[counter] = dist(bootstrap_counts, county_counts)
      counter = counter + 1
      #Stay tuned for the next episode
    }
  }
  
  return(rbind(rep(n_list, reps),dists))
}
```
## Experiment
We run our experiments.
```{r}
run_bootstraps(3, top_ten_counties_list)
```



